Supporting information: Copulas and their potential for ecology

### 2019-09-24 Paper.Rmd
editing results for Q1

<!--***DAN: You decided to reverse the decision to use pithy abbreviations for each
copula. Now the format is longer, so this might just annoy people. So this will take
some work to replace all those throughout the main text and sup mat,
including in figures and table and captions - everywhere.-->

In addition to our model selection approach, we also used a
nonparametric approach, to provide greater 
confidence in our answers to Q1. We used three 
statistics which quantify the extent to which 
the normalized ranks $u_i$ and $v_i$ are 
related in any part of their distributions; we here describe the 
statistics, with some additional details in Appendix \ref{SM-nonparam_stats}.
The statistics are defined with positively associated
variables in mind. All our variables were positively associated
when they were significantly associated.
Given two bounds 
$0 \leq l_b < u_b \leq 1$, define the lines $u+v=2l_b$ and 
$u+v=2u_b$, which intersect the unit square (Fig. \ref{NonparamStatsFig}). 
Our statistics quantify the association
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and $u_b\leq 0.5$ gives information about 
association in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ gives information about association in 
the right parts. The first statistic, a partial Spearman correlation
\begin{equation}
\cor_{l_b,u_b}(u,v)=\frac{\sum 
   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}},
\end{equation}
\noindent is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
between the bounds. 
Here sample means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$.
Larger values of the partial Spearman correlation indicate stronger positive association.
The sum of 
$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
interval $(0,1)$) equals the standard Spearman correlation
as long as no points fall directly on the bounds.
We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
which has a similar qualitative
interpretation to $\cor_{l_{b},u_{b}}$.
Our third statistic, $\Dsq_{l_b,u_b}$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$.
Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, for which large values inidicate strong association
between the bounds, small values of 
$\Dsq_{l_b,u_b}$ indicate strong association.

@@

For large datasets (large $n$),
we used $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we considered 
multiple bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$,
abbreviating $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.

To test for asymmetry of dependence in upper and lower portions of distributions,
we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
for smaller datasets (note the opposite order
in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ with $u_b$ close 
to $0$ for large datasets.
We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). Code and data for this
project are archived at \url{www.github.com/sghosh89/BIVAN}.

@@

<!--***DAN: Notes on what needs to be done to this section, so far only getting up through the presentation of the bivariate results:
1) be more nuanced in the presentation, don't give the appearance of sweeping results that dont agree with the general conclusions under the rug
2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text.
4) Sort the rows of Table S2 (and all analogous tables) AIC, putting the lowest
AIC at the top. This should mean the highest AICw is at the top.
5) move some of the content from appendices S6 and S7 into methods, and move fig S21 and table S4 into main text. These are non-parametric results for
soil C and N. So we will be presenting the nonparam results more comprehensively in the main text.
6) That means you also need to summarize the nonparm results for the other 3 bivariate systems in Table 2. How to do this? Maybe put the cor results only, for Spearman only (or Kendall, not sure which to choose) into Table 2.  
7) Probably ought to combine tables S6, S8, S10 into one table, and fix it so
whenever we currently cite one of those, change it so we cite the combined table instead.
8) The nonparam stats are not estimators of tail dependence, they measure 
something else (though related). Make it clear we know this. It turns out there are estimators
of tail dependence. Do we want to mention that?
-->

@@

Across essentially all our datasets, copula structures were typically non-normal and 
commonly showed asymmetric tail 
dependence, answering Q1 in the affirmative. 
To make results easier to absorb, we first 
present results for an example bivariate dataset, then we summarize results
for all bivariate datasets, then we present results for an example multivariate dataset, then we summarize for all multivariate datasets.

<!--bivariate example-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
normal copula had AIC weight essentially 0, answering the 
first part of Q1 for these data. 
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, answering the second part 
Q1 for these data. The values also differed substantially 
from each other, answering the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

Although these results convincingly show non-normal copula structure, 
we caution that the model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`). However, nonparametric statistics verified
both that tail dependence was distinct from that of a normal copula, and that it 
was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
Fig. \ref{fig_soilCN_nonparam}, Table \ref{tab_soilCN_asym}). Our 
affirmative, empirically based answer to Q1 for the soil C and N data is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.

@@

knitr::kable(h_ord, #longtable = T, 
             format = "latex",linesep = "",row.names = F, #digits = c(4,2,2,2,2,4,4), 
      caption.short = "Model fitting results for soil C and N dataset",
      caption = "Model fitting results for soil C and N dataset using 16 copula families: normal (N), Frank (F), Clayton (C), survival Clayton (SC), Gumbel (G), survival Gumbel (SG), Joe (J), survival Joe (SJ), BB1, survival BB1 (SBB1), BB6, survival BB6 (SBB6), BB7, survival BB7 (SBB7), BB8, survival BB8 (SBB8). Copulas are sorted according to their AIC values (2nd column). AICw = AIC weight; BIC = BIC value; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence. AIC, BIC and their corresponding weights are rounded upto 2 decimal place whereas tail dependences are rounded upto 4 decimal places. \\label{tab_soilCNfit}",
      booktabs = TRUE)
```

>>>>

In addition to our model selection approach, we also used a
nonparametric approach, to provide greater 
confidence in our answers to Q1. We used three 
statistics which quantify the extent to which 
the normalized ranks $u_i$ and $v_i$ are 
related in any part of their distributions; we here describe the 
statistics, with some additional details in Appendix \ref{SM-nonparam_stats}.
The statistics are defined with positively associated
variables in mind. All our variables were positively associated
when they were significantly associated.
Given two bounds 
$0 \leq l_b < u_b \leq 1$, define the lines $u+v=2l_b$ and 
$u+v=2u_b$, which intersect the unit square (Fig. \ref{NonparamStatsFig}). 
Our statistics quantify the association
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and $u_b\leq 0.5$ gives information about 
association in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ gives information about association in 
the right parts. The first statistic, a partial Spearman correlation
\begin{equation}
\cor_{l_b,u_b}(u,v)=\frac{\sum 
   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}},
\end{equation}
\noindent is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
between the bounds. 
Here sample means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$.
Larger values of the partial Spearman correlation indicate stronger positive association.
The sum of 
$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
interval $(0,1)$) equals the standard Spearman correlation
as long as no points fall directly on the bounds.
We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
which has a similar qualitative
interpretation to $\cor_{l_{b},u_{b}}$.
Our third statistic, $\Dsq_{l_b,u_b}$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$.
Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, for which large values inidicate strong association
between the bounds, small values of 
$\Dsq_{l_b,u_b}$ indicate strong association. These statistics are not estimators of the 
tail dependence quantities
defined previously, but rather are conceptually similar measures of associations in
the tail portions of the distributions when appropriate values of $l_b$ and $u_b$ are
selected.

@@

For large datasets (large $n$),
we used $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we considered 
multiple bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$,
abbreviating $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.
To test for asymmetry of dependence in upper and lower portions of distributions,
we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
for smaller datasets (note the opposite order
in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ with $u_b$ close 
to $0$ for large datasets.
We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also 
Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way to have no 
tail dependence (Appendix \ref{surrog_test}). 
The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the surrogate algorithm had two versions, one for preserving each correlation coefficient), 
but had normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were normal, but the data were otherwise 
statistically unchanged. The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on surrogate datasets 
provides a test of whether association between the variables in the part of the distributions specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure. 
Thus the comparison addresses the first two parts of Q1
for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower-tail associations 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper-tail associations different 
from that of a normal copula.
The comparison of one of the statistics 
$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
as computed on the empirical data, to the distribution of its values on many surrogate datasets 
provides a test of asymmetry of tail associations. Thus this comparison helps address the
third part of Q1 for bivariate 
datasets.

For multivariate datasets, we calculated Spearman and Kendall correlations
and the statistics 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
for all pairs of sampling locations, and we then computed means across the pairs
for each statistic. We used a spatial resampling scheme 
(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models.
Code and data for this
project are archived at \url{www.github.com/sghosh89/BIVAN}.

@@

<!--***DAN: Notes on what needs to be done to this section, so far only getting up through the presentation of the bivariate results:
1) be more nuanced in the presentation, don't give the appearance of sweeping results that dont agree with the general conclusions under the rug
DONE 2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text.
DONE 4) Sort the rows of Table S2 (and all analogous tables) AIC, putting the lowest
AIC at the top. This should mean the highest AICw is at the top.
DONE 5a) move some of the content from appendices S6 and S7 into methods, 
5b) and move fig S21 and table S4 into main text. These are non-parametric results for
soil C and N. So we will be presenting the nonparam results more comprehensively in the main text.
6) That means you also need to summarize the nonparm results for the other 3 bivariate systems in Table 2. How to do this? Maybe put the cor results only, for Spearman only (or Kendall, not sure which to choose) into Table 2.  
7) Probably ought to combine tables S6, S8, S10 into one table, and fix it so
whenever we currently cite one of those, change it so we cite the combined table instead.
DONE 8) The nonparam stats are not estimators of tail dependence, they measure 
something else (though related). Make it clear we know this. It turns out there are estimators
of tail dependence. Do we want to mention that?
-->

@@

Across datasets, copula structures were often non-normal and 
commonly showed asymmetric tail 
dependence, answering Q1 in the affirmative. 
To make results easier to absorb, we first 
present results for an example bivariate dataset, then we summarize results
for all bivariate datasets, then we present results for an example multivariate dataset, then we 
summarize for all multivariate datasets.

<!--bivariate example, model selection-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
normal copula had AIC weight essentially 0, answering the 
first part of Q1 for these data. 
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, helping to answer 
the second part 
Q1 for these data. The values also differed substantially 
from each other, helping answer the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
Our model selection procedures do not reveal whether the tail dependence values are 
*significantly* different from $0$ and from each other; @burnham2003_modelselection do
not recommend attempts to make statements about statistical significance of 
information-criterion-based, model selection results. 
Furthermore, we caution that, while our model selection results do convincingly show 
non-normal copula structure, model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`).
Our alternative nonparametric
approach, results detailed below, provides informaton about significance of tail
associations and asymmetries of tail associations.

<!--***DAN: Stopped here. Now do the nonpram results.-->

However, nonparametric statistics verified
both that tail dependence was distinct from that of a normal copula, and that it 
was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
Fig. \ref{fig_soilCN_nonparam}, Table \ref{tab_soilCN_asym}). Our 
affirmative, empirically based answer to Q1 for the soil C and N data is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.

@@

knitr::kable(h_ord, #longtable = T, 
             format = "latex",linesep = "",row.names = F, #digits = c(4,2,2,2,2,4,4), 
      caption.short = "Model fitting results for soil C and N dataset",
      caption = "Model fitting results for soil C and N dataset using 16 copula families: normal (N), Frank (F), Clayton (C), survival Clayton (SC), Gumbel (G), survival Gumbel (SG), Joe (J), survival Joe (SJ), BB1, survival BB1 (SBB1), BB6, survival BB6 (SBB6), BB7, survival BB7 (SBB7), BB8, and survival BB8 (SBB8). Table rows are sorted according by AIC. AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family with fitted parameters; UT = the same for upper-tail dependence. \\label{tab_soilCNfit}",
      booktabs = TRUE)
```

### 2019-09-24 SupportingInformation.Rmd
editing results for Q1

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way to have no 
tail dependence (Appendix \ref{surrog_test}). 
The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the surrogate algorithm had two versions, one for preserving each correlation coefficient), 
but had normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were normal, but the data were otherwise 
statistically unchanged. The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and similar 
Kendall (or Spearman)
correlation. Thus the comparison addresses the first two parts of Q1
for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.
The comparison of one of the statistics 
$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
as computed on the empirical data, to the distribution of its values on many surrogate datasets 
provides a test of asymmetry of dependence. Thus this comparison helps address the
third part of Q1 for bivariate 
datasets.

@@

For multivariate datasets, we calculated Spearman and Kendall correlations
and the statistics 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
for all pairs of sampling locations, and we then computed means across the pairs
for each statistic. We used a spatial resampling scheme 
(Appendix \ref{spatial_resamp}) to calculate confidence intervals of these means.

@@

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
normal-copula null hypothesis.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

>>>>

<!--***DAN: This appendix will eventually be eliminated when we move all the content to the 
main text. The next appendix (S7) will be retained pretty much as is. Probably S8 will all be moved to
the main text and S9 will be kept as it is.-->

@@

@@

For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
normal-copula null hypothesis.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].
